---
title: "Lab 3B: Manipulating, Analyzing, and Exporting Data with tidyverse "
author: "Alison Fowler"
output: html_document
---

The tidyverse package tries to address 3 common issues that arise when doing data analysis with some of the functions that come with R:

1. The results from a base R function sometimes depend on the type of data.

2. Using R expressions in a non standard way, which can be confusing for new learners.

3. Hidden arguments, having default operations that new learners are not aware of.

```{r echo = T, results = 'hide'}
library(tidyverse)
```

```{r}
# download data file from online 
download.file(url = "https://ndownloader.figshare.com/files/2292169",
              destfile = "data_raw/portal_data_joined.csv")

surveys <- read_csv("data_raw/portal_data_joined.csv")
# notice this is not read.csv(); read_csv() is in the readr package 
```


```{r}
# inspect the data 
str(surveys)
```


### Differences between a dataframe and a tibble: 

1. In addition to displaying the data type of each column under its name, it only prints the first few rows of data and only as many columns as fit on one screen.

2. Columns of class character are never converted into factors.


### Filter and select: 

```{r}
# select certain columns 
select(surveys, plot_id, species_id, weight)
```
```{r}
# select all columns except certain ones 
select(surveys, -record_id, -species_id)
```

```{r}
# choose rows based on specific criterion 
filter(surveys, year == 1995)
```

### Piping: 

What if you want to select and filter at the same time? 

There are three ways to do this: use intermediate steps, nested functions, or pipes.

```{r}
# Intermediate steps: 
surveys2 <- filter(surveys, weight < 5)
surveys_sml <- select(surveys2, species_id, sex, weight)

# Nest functions: 
surveys_sml <- select(filter(surveys, weight < 5), species_id, sex, weight)
```

These can be difficult to read... Pipes let you take the output of one function and send it directly to the next, which is useful when you need to do many things to the same dataset.

```{r}
surveys_sml <- surveys %>%
  filter(weight < 5) %>%
  select(species_id, sex, weight)

head(surveys_sml)
```

### Challenge: 

Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight. 

```{r}
surveys_sub <- surveys %>% 
  filter(year > 1995) %>% 
  select(year, sex, weight)

head(surveys_sub)
```

### Mutate 

allows us to create new columns based on the values in existing columns, for example to do unit conversions, or to find the ratio of values in two columns. 

```{r}
# create a new column of weight in kg: 

kg_added <- surveys %>% 
  mutate(weight_kg = weight/1000)

kg_and_lb_added <- surveys %>% 
  mutate(weight_kg = weight/1000, 
         weight_lb = weight_kg * 2.2)
```

```{r}
# if we wanted to remove NAs in weight column: 

surveys %>%
  filter(!is.na(weight)) %>%
  mutate(weight_kg = weight / 1000) %>%
  head()
```

### Challenge: 

Create a new data frame from the surveys data that meets the following criteria: contains only the species_id column and a new column called hindfoot_cm containing the hindfoot_length values converted to centimeters. In this hindfoot_cm column, there are no NAs and all values are less than 3. 

```{r}
new_HFL <- surveys %>% 
  mutate(hindfoot_cm = hindfoot_length / 10) %>% 
  filter(!is.na(hindfoot_cm)) %>% 
  filter(hindfoot_cm < 3) %>% 
  select(species_id, hindfoot_cm)

head(new_HFL)
```

### Split-apply-combine data analysis and the summarize() function

Split the data into groups, apply some analysis to each group, and then combine the results. dplyr makes this very easy through the use of the group_by() function.

The group_by() function is often used together with summarize(), which collapses each group into a single-row summary of that group. group_by() takes as arguments the column names that contain the categorical variables for which you want to calcuate the summary statistics. 

```{r}
# to compute the mean weight by sex: 

surveys %>% 
  group_by(sex) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE)) %>% 
  tail()
```

We can see that the sex column contains NA values because some animals had escaped before their sex and body weights could be determined. The resulting mean_weight column does not contain NA but NaN (which refers to “Not a Number”) because mean() was called on a vector of NA values whileat the same time setting na.rm = TRUE. To avoid this, we can remove the missing values for weight before we attempt to calculate the summary statistics on weight. Because the missing values are removed first, we can omit na.rm = TRUE when computing the mean:

```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight))
```

```{r}
# specify number of rows to display with print(n=)

surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight)) %>%
  print(n = 15)
```

Once the data are grouped, you can also summarize multiple variables at the same time (and not necessarily the same variable). For instance, we could add a column indicating the minimum weight for each species for each sex: 
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight))
```

```{r}
# We can sort on min_weight to put the lighter species first 

surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
  arrange(min_weight)

# and we can sort in descending order... 

surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
  arrange(desc(mean_weight)) 
```

### Counting

We often want to know the number of observations found for each factor or combination of factors. For this task, dplyr provides count(). 

```{r}
# if we wanted to count the number of rows of data for each sex: 

surveys %>% 
  count(sex)
```

```{r}
# which is equivalent to: 

surveys %>% 
  group_by(sex) %>% 
  summarize(count = n())
```

```{r}
# there is a sort argument for count() 

surveys %>% 
  count(sex, sort = T)
```

```{r}
# multiple factors: specify the order 

surveys %>%
  count(sex, species)

surveys %>%
  count(species, sex)
```

```{r}
# arrange species in alphabetical order and count in descending order 

surveys %>%
  count(sex, species) %>%
  arrange(species, desc(n))
```

### Challenge: 

1. How many animals were caught in each plot_type surveyed?

```{r}
surveys %>% 
  count(plot_type)
```

2. Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).

```{r}
spp_HFL <- surveys %>% 
  group_by(species_id) %>% 
  summarize(mean_length = mean(hindfoot_length),
            min_length = min(hindfoot_length),
            max_length = max(hindfoot_length))

head(spp_HFL)

# Oh, it can't calculate means with NAs; let's remove...  

spp_HFL <- surveys %>% 
  filter(!is.na(hindfoot_length)) %>%
  group_by(species_id) %>% 
  summarize(mean_length = mean(hindfoot_length),
            min_length = min(hindfoot_length),
            max_length = max(hindfoot_length),
            n = n())

head(spp_HFL)
```

3. What is the heaviest animal measured each year? Return the columns year, genus, species_id, and weight 

```{r}
spp_weights <- surveys %>% 
  filter(!is.na(weight)) %>%
  group_by(year) %>% 
  filter(weight == max(weight)) %>% # filter, not summarize! 
  select(year, species_id, genus, weight) %>% 
  arrange(year)

head(spp_weights)
```

### Reshaping with gather and spread 

spread() takes three principal arguments: 
1. the data
2. the key column variable whose values will become new column names
3. the value column whose values will fill the new column variables 

further arguments include fill which, if set, fills in missing values with the value provided 

use spread() to transform surveys to find the mean weight of each genus in each plot over the entire period. 

```{r}
surveys_gw <- surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(plot_id, genus) %>% 
  summarize(mean_weight = mean(weight))

head(surveys_gw)
str(surveys_gw)
```

Observations for each plot are spread across multiple rows, 196 obs of three variabes. 

```{r}
surveys_spread <- surveys_gw %>% 
  spread(key = genus, value = mean_weight)

head(surveys_spread)
str(surveys_spread)

surveys_gw %>%
  spread(genus, mean_weight, fill = 0) %>%
  head()
```

gather() can be used to do the opposite; takes in four principle arguments: 
1. the data
2. the key column variable we wish to create from column names.
#. the values column variable we wish to create and fill with values associated with the key.
4. the names of the columns we use to fill the key variable (or to drop).

```{r}
surveys_gather <- surveys_spread %>%
  gather(key = "genus", value = "mean_weight", -plot_id)

head(surveys_gather)
str(surveys_gather)
```

```{r}
# you can also specify what columns to include - but for some reason when I tried to change the range of genera, it got messed up
surveys_spread %>%
  gather(key = "genus", value = "mean_weight", Baiomys:Spermophilus) 
```

### Challenge: 

1. Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It’s a powerful function! See ?n_distinct for more.

```{r}
genera_per_plot <- surveys %>% 
  group_by(plot_id, year) %>% 
  summarize(genera = n_distinct(genus))
  # I guess it knows to do distinct per plot_id and not year because it is listed first? 

# now lets add spread 

genera_per_plot <- surveys %>% 
  group_by(plot_id, year) %>% 
  summarize(genera = n_distinct(genus)) %>%
  spread(year, genera)

head(genera_per_plot)
```

2. Now take that data frame and gather() it again, so each row is a unique plot_id by year combination. 

```{r}
gathered_genera <- genera_per_plot %>% 
  gather("year", "genera", -plot_id) 
# why does gather need columns to be in quotes? 

head(gathered_genera)
```

3. The surveys data set has two measurement columns: hindfoot_length and weight. This makes it difficult to do things like look at the relationship between mean values of each measurement per year in different plot types. Let’s walk through a common solution for this type of problem. 

First, use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight. Hint: You’ll need to specify which columns are being gathered.

```{r}
surveys_long <- surveys %>% 
  gather ("measurement", "value", hindfoot_length, weight)

str(surveys_long)
```

This doubled the amount of rows. It added hindfootlength to one and then weight to another for each obs.

4. With this new data set, calculate the average of each measurement in each year for each different plot_type. Then spread() them into a data set with a column for hindfoot_length and weight. Hint: You only need to specify the key and value columns for spread()

```{r}
measurements <- surveys_long %>% 
  group_by(year, measurement, plot_type) %>% 
  summarize(mean_value = mean(value, na.rm=T)) %>% 
  spread(measurement, mean_value)

head(measurements)
```

### Exporting Data 

First, remove Nas 

```{r}
surveys_complete <- surveys %>% 
  filter(!is.na(weight),
         !is.na(hindfoot_length),
         !is.na(sex))
```

Remove rare species; first make a subset of the most common species (50 or more observations)

```{r}
species_counts <- surveys_complete %>%
  count(species_id) %>% 
  filter(n >= 50)
```

Then subset the complete data set with those species names 

```{r}
surveys_complete <- surveys_complete %>% 
  filter(species_id %in% species_counts$species_id)
```

Double check we have the right dimensions (30,463 x 13)

```{r}
dim(surveys_complete)
```

Write new csv file 

```{r}
write_csv(surveys_complete, path = "generated_data/surveys_complete.csv")
```



