---
title: "Lab 5 - COVID data Part 2"
author: "Alison Fowler"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# Read in data 
# March 13
Confirmed_State_3_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State") %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 

str(Confirmed_State_3_13)
```

```{r}
# Sept 13
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 

str(Confirmed_State_9_13)
```

```{r}
# Notice from the above `str’ calls that there are 53 states in the table on 3/13/2020 and 
# 58 states in the table on 9/13/2020... Let's see what the differences are 

setdiff(Confirmed_State_9_13$Province_State, Confirmed_State_3_13$Province_State)
```

```{r}
# There are 4 US territories and a field you may have notice last week: “Recovered”. 
# We don’t need the “Recovered” row so let’s filter it from the data set

Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered") 
```

```{r}
# Since all of the states and territories in 9/13/2020 are found in 3/13/2020 a left_join or full_join can 
# be used to merge the tables

Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
                                              Confirmed_State_9_13, by = c("Province_State"))
```

```{r}
# look at the new joined dataframe 
head(Confirmed_State_3_13_9_13_joined)
tail(Confirmed_State_3_13_9_13_joined, 5)
```

```{r}
# look to see what cells are NA  
which(is.na(Confirmed_State_3_13_9_13_joined))
```

```{r}
# let's turn the NAs to 0 and rename the columns 
Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
                                              Confirmed_State_9_13, by = c("Province_State")) %>% 
  rename(Confirmed_3_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
  select(-Country_Region.x, -Country_Region.y) %>% 
  replace_na(list(Confirmed_3_13_2020 = 0))

head(Confirmed_State_3_13_9_13_joined)
```

```{r}
# Switching between wide and long table formats
Confirmed_State_3_13_9_13_joined_long <- Confirmed_State_3_13_9_13_joined %>% 
  pivot_longer(-c(Province_State),
               names_to = "Date", values_to = "Confirmed")
```

```{r}
# to adjust figure size {r, fig.width = 5, fig.height = 10}
ggplot(Confirmed_State_3_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + 
  geom_point(aes(color = Date)) 
```

```{r}
# Working with the time series data 
# load the most recent time series data for confirmed cases 

time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", 
              destfile = "time_series_covid19_confirmed_global.csv")

time_series_confirmed <- read_csv("time_series_covid19_confirmed_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

head(time_series_confirmed)
```

```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Confirmed") 

head(time_series_confirmed_long)
```

```{r}
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", 
              destfile = "time_series_covid19_deaths_global.csv")

time_series_deaths <- read_csv("time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths_long <- time_series_deaths %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths") 

head(time_series_deaths_long)
```

```{r}
# Joining the time series tables 

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)

head(time_series_confirmed_long)
```

```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

head(time_series_deaths_long)
```

```{r}
# now join the tables 
time_series_long_joined <- full_join(time_series_confirmed_long,
                                     time_series_deaths_long, by = c("Key")) %>% 
  select(-Key)

head(time_series_long_joined)
```

```{r}
which(is.na(time_series_long_joined$Confirmed))
```

```{r}
which(is.na(time_series_long_joined$Deaths))
```

# no NAs - great 

```{r}
library(lubridate)
```

```{r}

```

```{r}

```

```{r}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)

# create a column using pivot_long 

time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
```

```{r}
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")

head(time_series_long_joined_counts)
```

```{r}
# Making graphs from the time series data 

time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
  ggplot(aes(x = Date,  y = Deaths)) + 
  geom_point() +
  geom_line() +
  ggtitle("US COVID-19 Deaths")
```

```{r}
# some other countries 
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region %in% c("China","Japan", "Korea, South",
                                "Italy","Spain", "US")) %>% 
  ggplot(aes(x = Date,  y = Deaths)) + 
  geom_point() +
  geom_line() +
  ggtitle("COVID-19 Deaths") +
  facet_wrap(~Country_Region, ncol=2, scales="free_y")
```

```{r}
# several countries on the same graph

time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region %in% c("China","France","Italy", 
                                "Korea, South", "US")) %>% 
  ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
  geom_point() +
  geom_line() +
  ggtitle("COVID-19 Deaths")
```

```{r}
# cases and deaths on the same graph 

time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
  ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
  geom_point() +
  geom_line() +
  ggtitle("US COVID-19 Cases")
```

### Exercises 
1. Finish the visualization tutorial (Lab 5A)

2. Revise the above example for joining tables using 6/13/2020 and 9/13/2020. Plot the data as a bar plot. Make sure your graph looks nice in the report by adjusting the height in the R code chunk header.

Add more informative x and y labels and a title to the above graph in Ex2.

Using the time series data. Make a plot of the total number of confirmed deaths per day worldwide

Use Mutate to make a new column with deaths/confirmed cases (Hint: this might be best done after joining the tables, but before pivoting longer).

Plot US deaths/confirmed cases per day.

Make a single graph with from the times series data with the 10 countries with the hightest death total.

Using Facet_Wraps make a graph from the times series data with the 10 countries with the hightest death total. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

Using Facet_Wraps make a graph from the times series data for confirmed cases for all US states and territories. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

Apply a theme other than the one used in the Data Carpentries tutorial to the graph in Ex8.



