---
title: "Lab 5A: Visualization with ggplot"
output: html_document
---

```{r echo = FALSE}
library(ggplot2)
library(tidyverse)
```

```{r echo = FALSE}
surveys_complete <- read_csv("generated_data/surveys_complete.csv")
```

Boxplots 
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_boxplot()
```

Add points 
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_boxplot(alpha = 0) +
  geom_jitter(alpha = 0.3, color = "tomato")
```

Put the boxes in front of the points... 
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato") +
  geom_boxplot(alpha = 0)
```

### Challenges 

Try a violin plot
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato") +
  geom_violin(alpha = 0)
```

Without the points...
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_violin(alpha = 0)
```

Change scale
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_violin(alpha = 0) + 
  scale_y_log10()
```

Plot hindfoot length 
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
  geom_boxplot(alpha = 0) 
```

Change plot_id from integer to factor so that it's not a continuous variable 
```{r}
surveys_complete[,'plot_id'] <- factor(surveys_complete[,'plot_id'])
class(surveys_complete$plot_id)
```

Plot again - not working ?? need to fix
```{r}
#ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length, color = plot_id)) +
#  geom_jitter() +
#  geom_boxplot() 
```

Hmm for some reason it doesn't like the color line 

### Plotting Time Series Data

Counts per year per genus 
```{r}
yearly_counts <- surveys_complete %>%
  count(year, genus)
```

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
  geom_line()
```

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, color = genus)) +
  geom_line()
```

Integrating pipe operator with ggplot
```{r}
yearly_counts %>% 
  ggplot(mapping = aes(x = year, y = n, color = genus)) +
  geom_line()
```

Link data manipulation and visualization
```{r}
yearly_counts_graph <- surveys_complete %>%
  count(year, genus) %>% 
  ggplot(mapping = aes(x = year, y = n, color = genus)) +
  geom_line()

yearly_counts_graph
```

Faceting
```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
  geom_line() +
  facet_wrap(facets = vars(genus))
```

Split the sexes 
```{r}
yearly_sex_counts <- surveys_complete %>%
  count(year, genus, sex)
```

```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(facets =  vars(genus))
```

Split sexes and genera

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```

One column, facet by rows
```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))
```

One row, facet by column
```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols = vars(genus))
```

### ggplot2 themes 
https://ggplot2.tidyverse.org/reference/ggtheme.html 
```{r}
library(ggthemes)
```

Change the background to white 
```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(vars(genus)) +
  theme_bw()
```

### Challenge 
Create a plot that depicts how the average weight of each species changes through the years 

```{r}
yearly_weights_graph <- surveys_complete %>%
  group_by(year, genus) %>% 
  summarize(mean_weight = mean(weight)) %>% 
  ggplot(mapping = aes(x = year, y = mean_weight, color = genus)) +
  geom_line()

yearly_weights_graph 
```

### Customization 

Add axis titles 
```{r}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(vars(genus)) +
  labs(title = "Observed genera through time",
       x = "Year of observation",
       y = "Number of individuals") +
  theme_bw()
```

Increase axis titles font size 
```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(vars(genus)) +
  labs(title = "Observed genera through time",
       x = "Year of observation",
       y = "Number of individuals") +
  theme_bw() +
  theme(text=element_text(size = 16))
```

Change orientation of the labels & italicize 
```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(vars(genus)) +
  labs(title = "Observed genera through time",
       x = "Year of observation",
       y = "Number of individuals") +
  theme_bw() +
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 12),
        strip.text = element_text(face = "italic"),
        text = element_text(size = 16))
```

Save your changes as a theme
```{r}
grey_theme <- theme(axis.text.x = element_text(colour="grey20", size = 12, 
                                               angle = 90, hjust = 0.5, 
                                               vjust = 0.5),
                    axis.text.y = element_text(colour = "grey20", size = 12),
                    text=element_text(size = 16))

ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length)) +
  geom_boxplot() +
  grey_theme
```

### Challenge 
Change one of the graphs to make it pretty 

We'll look at species counts in the different plot types 
```{r}
class(surveys_complete$plot_type)
surveys_complete$plot_type<-as.factor(surveys_complete$plot_type)
levels(surveys_complete$plot_type)

yearly_plot_types<-surveys_complete %>% 
  count(year, genus, plot_type)

head(yearly_plot_types)
```

```{r}
ggplot(data = yearly_plot_types, mapping = aes(x = year, y = n, color = plot_type)) +
  geom_line(size = 1) + # thicker lines
  facet_wrap(vars(genus)) +
  labs(title = "Observed genera through time",
       x = "Year of observation",
       y = "Number of individuals") +
  theme_bw() +
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 12),
        strip.text = element_text(face = "italic"),
        text = element_text(size = 16)) + 
  scale_fill_discrete(name = "Plot Type") # why isn't this working? 
```

### Arranging and exporting plots 

Put multiple plots together

```{r echo = FALSE}
library(gridExtra)
```

```{r}
spp_weight_boxplot <- ggplot(data = surveys_complete, 
                             aes(x = species_id, y = weight)) +
  geom_boxplot() +
  labs(x = "Species", 
       y = expression(log[10](Weight))) +
  scale_y_log10() + 
  labs()

spp_count_plot <- ggplot(data = yearly_counts, 
                         aes(x = year, y = n, color = genus)) +
  geom_line() + 
  labs(x = "Year", y = "Abundance")

grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, widths = c(4, 6))
```

Export 
```{r}
my_plot <- ggplot(data = yearly_sex_counts, 
                  aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(vars(genus)) +
  labs(title = "Observed genera through time",
       x = "Year of observation",
       y = "Number of individuals") +
  theme_bw() +
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 12),
        text = element_text(size = 16))

ggsave("name_of_file.png", my_plot, width = 15, height = 10)
```

This also works for grid.arrange() plots

```{r}
combo_plot <- grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, 
                           widths = c(4, 6))
ggsave("combo_plot_abun_weight.png", combo_plot, width = 10, dpi = 300)
```

