---
title: "Lab 9 extra : Programming in R - Control Structures and Functions"
author: "Alison Fowler"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Control Structures 

Commonly used control structures are: 

* if and else: testing a condition and acting on it
* for: execute a loop a fixed number of times
* while: execute a loop *while* a condition is true

**Conditional statements with if-else**

```{r, eval = FALSE}
if(<condition>) {
        ## do something
} 
## Continue with rest of code
```
 
```{r, eval = FALSE}
if(<condition>) {
        ## do something
} 
else {
        ## do something else
}
```


```{r, eval = FALSE}
if(<condition1>) {
        ## do something
} else if(<condition2>)  {
        ## do something different
} else {
        ## do something different
}
```


```{r}
## Generate a uniform random number
x <- runif(1, 0, 10)  

x

if(x > 3) {
        y <- 10
} else {
        y <- 0
}

y
```

**Important Operators**

Comparison operators:

* == (equal)
* != (not equal)
* ">" (greater than)
* ">=" (greater than or equal)
* < (less than)
* <= (less than or equal)

Logical operators:

* & (and)
* | (or)
* ! (not)

**for Loops**

```{r}
for(i in 1:5) {
  print (i)
  }
```

```{r}
x <- c("a", "b", "c", "d")

for(i in 1:4) {
        ## Print out each element of 'x'
        print(x[i])  
}
```

```{r}
## Generate a sequence based on length of 'x'
for(i in seq_along(x)) {   
        print(x[i])
}
```


```{r}
for(letter in x) {
        print(letter)
}
```

```{r}
for(i in 1:4) print(x[i])
```

**Nested for Loops**

```{r}
x <- matrix(1:6, 2, 3)
for(i in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[i, j])
        }   
}
```

**while Loops**

```{r}
count <- 0
while(count < 10) {
        print(count)
        count <- count + 1
}
```


```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  ## random walk
                z <- z + 1
        } else {
                z <- z - 1
        } 
}
print(z)
```


### Functions 

Here's a simple function that takes no arguments and does nothing.
```{r}
f <- function() {
         ## This is an empty function
}

class(f)
```

```{r}
f()
```

Now let's create a function that actually has a function body

```{r}
f <- function() {
  cat("Hello, world!\n")
}

f()
```

```{r}
f <- function(num) {
        for(i in seq_len(num)) {
                cat("Hello, world!\n")
        }
}

f(3)
```

```{r}
f <- function(num) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
               cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
meaningoflife <- f(3)

print(meaningoflife)
```


```{r, eval = FALSE}
f()
```

```{r}
f <- function(num = 1) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}

f()    ## Use default value for 'num'
```

```{r}
f(2)   ## Use user-specified value
```

```{r}
f(num = 2)
```


